To demonstrate taking a simple random sample (SRS), we'll use the `mtcars` dataset built into R:

```{r}
head(mtcars)
```

Our goal here is to randomly sample 5 rows (i.e., car models) from the full `mtcars` dataset.

The easiest way to create an SRS in `R` is using the special function `sample_n()` which is found in the `dplyr` package. `R` "packages" are collections of useful tools that augment the functionality provided by the basic version of `R`. Packages are always free and are downloaded from the internet.

To download, install, and load the `dplyr` package:

1. Make sure you have an internet connection.
2. Select the "Packages" tab from the bottom-right-hand pane in RStudio, and click on "Install".
3. In the "Install From" dropdown, select "Repository (CRAN, CRANextra)". CRAN is a website which stores R packages.
4. Below, in the "Packages" box, type in `dplyr` (it should auto-complete).
5. Make sure "Install dependancies" is checked and click "Install". You may be prompted to choose a CRAN mirror. Choose one of the USA options.
6. RStudio will download, configure, and install the package automatically; you will see some gibberish in your console.
7. Load the newly downloaded package into your workspace by typing the following code into your console:
    ```{r, warning=F, message=F}
    library(dplyr)
    ```

You now have access to the `sample_n()` function! You can look at the manual for this function by typing the following into your console:
```{r, eval=F}
?sample_n
```
This will pop up some information about the function in the "Help" tab of the bottom-left pane in RStudio.

We're finally ready to take an SRS from the `mtcars` dataset. We need to give `sample_n()` two pieces of information: the name of the dataset we want to sample, and the size of the sample (e.g., 5 subjects). We will save our new sample into a new object called `sample1`.

```{r}
sample1 <- sample_n(mtcars, size = 5)
```

View the saved `sample1` object in the console. Note yours will not be exactly the same set of rows as what you see here, it is a random sample after all!

```{r}
sample1
```

To get another SRS from `mtcars`, simply run the same command from above again, but save to a different object (e.g. `sample2`):

```{r}
sample2 <- sample_n(mtcars, size = 5)
sample2
```
