### Quantitative Data

R has many built in tools to calculate descriptive statistics of quantitative data. We'll calculate a series of descriptive statistics for the `mpg` (miles-per-gallon) column of the `mtcars` dataset that comes with R. To review, here's what the first few rows of the `mtcars` dataset look like:

```{r}
head(mtcars)
```

First, we'll save the `mpg` column into a new object called `mydata`, using the `$` operator (see the end of the "Importing Data" section of the manual for more details)

```{r}
mydata <- mtcars$mpg
mydata
```

Next, we can calculate the mean (or average) value of mpg using the `mean()` function, which takes the data object as an input and outputs the corresponding statistic into the console:

```{r}
mean(mydata)
```

There are many other built-in R functions that you will find useful. For example:

```{r}
median(mydata)
sd(mydata)
range(mydata)
```

Giving you the median, standard deviation, and range of the dataset, respectively. You can also get the five-number summary of your dataset (plus the mean):

```{r}
summary(mydata)
```

R can also be used to calculate how many values in a dataset fall in a certain range. For example, say you wanted to know how many values in `mydata` were less than 2. The following code checks each individual element of `mydata` to see if it is less than (`<`) 2; if it is, it returns `TRUE`, and if not `FALSE`.

```{r}
mydata < 21
```

The next step is to call the `sum` function on the output, which will total up the number of `TRUE` entries and thus the number of values less than 2:

```{r}
sum(mydata < 21)
```

You can do much more complicated checks as well. For example, say you wanted to know how many elements of `mydata` were *less than or equal to* 2 AND *greater than* 1. You can combine two checks with `&`:

```{r}
sum(mydata <= 21 & mydata > 14)
```

To calculate the percentage of values of `mydata` less than or equal to 2 and greater than 1, simply divide the above code by the length of mydata and multiply by 100:

```{r}
sum(mydata <= 21 & mydata > 14) / length(mydata) * 100
```

### Categorical Data

The `mtcars` dataset also contains a column `cyl` which contains information about the number of cylinders (4, 6, or 8) in the engine of each car model. Although `cyl` contains numbers, we can treat it as a categorical variable. 

Create a new object called `cyldata` that contains only the `cyl` column, using the `$` operator:

```{r}
cyldata <- mtcars$cyl
```

The easiest way to summarize categorical data is to create a table counting up the number of car models in each cylinder category, using the `table()` function:

```{r}
table(cyldata)
```

Generating a table like this one will also be useful when creating barplots and pie charts (see Graphs and Plots section).
